Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields - I

### 2016-10-04 ms_disp1/ms_disp1.tex
new figures

\usepackage{graphicx,longtable,natbib,mathrsfs,color}
\usepackage{txfonts}

@@

%
\def\apjl{Astrophys. J. Lett.}
\def\apjs{Astrophys. J. Suppl. Ser.}
\def\aj{Astron. J.}
\def\mnras{Mon. Not. R. Astron. Soc.}
\def\aap{Astron. Astrophys.}
\def\jcap{J. Cosmology Astropart. Phys.}
\def\aapr{Astron. Astrophys. Rev.}

@@

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}

@@

\maketitle

\section{Introduction}\label{sec.intro}
Our universe starts from primordial Gaussian perturbations at a very early stage, 
and from those fluctuations, the gravitational instability drives the formation of 
the large scale structure (LSS) distribution of matter
\citep{1970A&A.....5...84Z,1985ApJ...292..371D}.
These structures grow 
linearly until the perturbations are large enough so that the first order 
perturbation theories are unable to analytically describe the LSS distributions 
\citep{2016JCAP...01..043M}. 
As a result, the final nonlinear LSS distribution contains higher order 
statistics, and thus makes it more challenging to be interpreted into basic
cosmological parameters. One such example is that, the baryonic acoustic oscillation (BAO)
scale can be used as a ``standard ruler'' to constrain the cosmic expansion history
and thus probes the dark energy properties \citep{2005NewAR..49..360E}, but
nonlinear evolution smears the BAO features and lowers the measurement
accuracy \citep{2005ApJ...633..560E,2012MNRAS.419.2949N}.
There are various attempts to recover earlier stages
of LSS, in which statistics are closer to Gaussian 
\citep{1992MNRAS.254..315W,2013MNRAS.436..759H}.
Because Gaussian fields can be adequately described by two-point statistics,
ideally after some recovery algorithms, more information can be extracted,
more straightforwardly, by power spectra or two-point correlation functions
\citep{2005MNRAS.360L..82R,2012MNRAS.421..832Y}.

Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement
\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field
$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
Of course, the full displacement field $\bs\Psi(\bs q)$
is non-observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
\tcr{[Peacock, Goldberg?]}
In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
simplifies to an ordering of matter elements by Eulerian coordinates.
Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
displacement field $s(q)$, and find that \tcr{this new method} well recovers the linear
information and reconstructs BAO peak in correlation function. 
In 3D case, there are various techniques to obtain $\bs s(\bs q)$.
However one needs to carefully consider effects of curl, shell-crossing in 3D case.

@@

\begin{figure}[b] \centering
  \includegraphics[width=1.0\linewidth]{fig1.pdf}
  \caption{Dimensionless auto- and cross- power spectra between nonlinear density field $\delta$, linear density field $\delta_L$ and reconstructed density field $\delta_R$.}
  \label{fig.1}
\end{figure}
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$
  (solid lines). Their
  damping of the power spectra $W^2$ (dashed lines),
  and their best fit BAO damping models (dotted lines).}
  \label{fig.2}
\end{figure}

@@

\subsection{Reconstruction}\label{ss.reco}
In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
Because the reconstruction processes are implemented on Lagrangian coordinates,
$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
We just write $\bs q$'s Fourier wave number
$k_q$ as $k$ to simplify the expression.
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
up to a relatively smaller scale (higher $k$).
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig3.pdf}
  \caption{Power spectra of the nonlinear density field and its
  filtered density field.}
  \label{fig.3}
\end{figure}

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig4.pdf}
  \caption{Power spectra of the reconstructed density field and its
  filtered density field.}
  \label{fig.4}
\end{figure}

>>>>

\usepackage{graphicx,longtable,natbib,mathrsfs,color}
\usepackage{txfonts,aas_macros}

@@

@@

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields - I}

@@

\maketitle
\begin{figure}
 \centering
  \includegraphics[width=0.99\linewidth]{denx3.pdf}
  \caption{}
  \label{fig.1}
\end{figure}
\section{Introduction}\label{sec.intro}
Our universe starts from primordial Gaussian perturbations at a very early stage, 
and from those fluctuations, the gravitational instability drives the formation of 
the large scale structure (LSS) distribution of matter
\citep{1970A&A.....5...84Z,1985ApJ...292..371D}.
These structures grow 
linearly until the perturbations are large enough so that the first order 
perturbation theories are unable to analytically describe the LSS distributions 
\citep{2016JCAP...01..043M}. 
As a result, the final nonlinear LSS distribution contains higher order 
statistics, and thus makes it more challenging to be interpreted into basic
cosmological parameters. One such example is that, the baryonic acoustic oscillation (BAO)
scale can be used as a ``standard ruler'' to constrain the cosmic expansion history
and thus probes the dark energy properties \citep{2005NewAR..49..360E}, but
nonlinear evolution smears the BAO features and lowers the measurement
accuracy \citep{2005ApJ...633..560E,2012MNRAS.419.2949N}.
There are various attempts to recover earlier stages
of LSS, in which statistics are closer to Gaussian 
\citep{1992MNRAS.254..315W,2013MNRAS.436..759H}.
Because Gaussian fields can be adequately described by two-point statistics,
ideally after some recovery algorithms, more information can be extracted,
more straightforwardly, by power spectra or two-point correlation functions
\citep{2005MNRAS.360L..82R,2012MNRAS.421..832Y}.

Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement
\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field
$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
Of course, the full displacement field $\bs\Psi(\bs q)$
is non-observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
\tcr{[methods of Peebles, david Goldberg? least action ]}
In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
simplifies to an ordering of matter elements by Eulerian coordinates.
Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
displacement field $s(q)$, and find that \tcr{this new method} well recovers the linear
information and reconstructs BAO peak in correlation function. 
In 3D case, there are various techniques to obtain $\bs s(\bs q)$.
However one needs to carefully consider effects of curl, shell-crossing in 3D case.

@@

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{fig2.pdf}
  \caption{Correlation functions $r(\delta_L,\delta)$ and $r(\delta_L,\delta_R)$
  (solid lines). Their
  damping of the power spectra $W^2$ (dashed lines),
  and their best fit BAO damping models (dotted lines).}
  \label{fig.2}
\end{figure}

@@

\subsection{Reconstruction}\label{ss.reco}
In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
Because the reconstruction processes are implemented on Lagrangian coordinates,
$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
We just write $\bs q$'s Fourier wave number
$k_q$ as $k$ to simplify the expression.
In Fig.\ref{fig.1} we show the power spectrum of $\delta_R$ and its cross power
with $\delta_L$. Despite of a lowered power of $\delta_R$ compared to $\delta$,
it has a much higher cross power with $\delta_L$ compared to $\delta_N$,
up to a relatively smaller scale (higher $k$).
\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{recopower.pdf}
  \caption{Power spectra of the nonlinear density field and its
  filtered density field.}
  \label{fig.3}
\end{figure}

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{pdfs.pdf}
  \caption{Power spectra of the reconstructed density field and its
  filtered density field.}
  \label{fig.4}
\end{figure}
